simodels: A package for Spatial Interaction Modelling

### 2024-08-28 paper/paper.qmd
Update paper.qmd in conversation with @SebnemEr

@@

```{r validation}
od_dfjc = inner_join(od_dfj |> select(-all), od)
od_dfjc |> 
  ggplot() +
  geom_point(aes(all, flow_constrained))
cor(od_dfjc$all, od_dfjc$flow_constrained)^2
```

>>>>

```{r validation}
# od_dfjc = inner_join(od_dfj |> select(-all), od)
# od_dfjc |> 
#   ggplot() +
#   geom_point(aes(all, flow_constrained))
# cor(od_dfjc$all, od_dfjc$flow_constrained)^2
```

### Commuter flows in Leeds, UK

See https://github.com/acteng/netgen/blob/main/odgen.md for ideas on this with pupils data.

And https://github.com/acteng/netgen/pull/10/files

```{r}
od_observed = simodels::si_oa_wpz
dim(od_observed)
zones = simodels::si_oa_wpz_o
destinations = simodels::si_oa_wpz_d
```

```{r}
od_modelled_max_5km = si_to_od(zones, destinations)
od_modelled = si_to_od(zones, destinations, max_dist = 5000)
dim(od_modelled)
names(od_modelled)
```

```{r}
#| label: simple-model
#| echo: true
gravity_model = function(beta, d, m, n) {
  m * n * exp(-beta * d / 1000)
} 
# perform SIM
od_res = simodels::si_calculate(
  od_modelled,
  fun = gravity_model,
  d = distance_euclidean,
  m = origin_n_o,
  n = destination_n_d,
#   constraint_production = origin_all,
  beta = 0.8
  )
od_res_df = od_res |>
  sf::st_drop_geometry()
```

```{r}
names(od_observed)
od_joined = left_join(
  od_res_df,
  od_observed |>
    select(O = OA11CD, D = wz11cd, n_observed = n)
)
```

```{r}
#| label: plot_od_fit
#| include: false
# Aim: create function that takes in od_res and returns a ggplot object
plot_od_fit = function(od_res, title = "(unconstrained)") {
  res_o = od_res |>
    group_by(O) |>
    summarise(
      Observed = first(origin_n_o),
      Modelled = sum(interaction),
      Type = "Origin"
    )
  res_d = od_res |>
    group_by(D) |>
    summarise(
      Observed = first(destination_n_d),
      Modelled = sum(interaction),
      Type = "Destination"
    )
  res_od = od_res |>
    transmute(
      Observed = n_observed,
      Modelled = interaction,
      Type = "OD"
    )
  res_combined = bind_rows(res_o, res_d, res_od) |>
    # Create ordered factor with types:
    mutate(
      Type = factor(Type, levels = c("Origin", "Destination", "OD"))
    )
  rsq_o = cor(res_o$Observed, res_o$Modelled, use = "complete.obs")^2
  rsq_d = cor(res_d$Observed, res_d$Modelled, use = "complete.obs")^2
  rsq_od = cor(res_od$Observed, res_od$Modelled, use = "complete.obs")^2
  rsq_summary = data.frame(
    rsq = c(rsq_o, rsq_d, rsq_od) |> round(3),
    Type = c("Origin", "Destination", "OD")
  )
  g_combined = res_combined |>
    left_join(rsq_summary) |>
    # Add rsquared info:
    mutate(Type = paste0(Type, " (R-squared: ", rsq, ")")) |>
    # Update factor so it's ordered (reverse order):
    mutate(
      Type = factor(Type, levels = unique(Type))
    ) |>
    ggplot() +
      geom_point(aes(x = Observed, y = Modelled)) +
      geom_smooth(aes(x = Observed, y = Modelled), method = "lm") +
      facet_wrap(~Type, scales = "free") +
      labs(
        title = paste0("Model fit at origin, destination and OD levels ", title),
        x = "Observed",
        y = "Modelled"
      )
  g_combined
}
# Test it:
g = plot_od_fit(od_joined)
```

```{r}
#| label: plot-od-fit
#| fig-cap: Model fit at origin, destination and OD levels
g
```

### Travel to school in York, UK

@@
